Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-06 paper/paper.tex
revise the code

\begin{lstlisting}
type Addr = Int
type Env = Map[String, Addr]
type Store = Map[Addr, Storable]

abstract class Storable
case class Clos(v: Lam, env: Env) extends Storable

case class Frame(x: String, e: Expr, env: Env)
type Kont = List[Frame]

case class State(e: Expr, env: Env, store: Store, k: Kont)
\end{lstlisting}

@@

\begin{lstlisting}
abstract class Storable
case class Clos(v: Lam, env: Env) extends Storable

case class Frame(x: String, e: Expr, env: Env)
case class Cont(frame: Frame, kaddr: KAddr)
\end{lstlisting}

@@

In the definition of @State@, the continuation store disappears, and
component @konts@ becomes a list of frames. 
The other components remain unchanged as in the Section \ref{aam}.

@@

\begin{lstlisting}
case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
case class NDState(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time, ndk: List[NDCont]) {
  def toState: State = State(e, env, bstore, konts, time)
}
\end{lstlisting}

@@

\begin{lstlisting}
case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
  konts match {
    case Nil => ndk match {
      case NDCont(Nil, _, _, _, _)::ndk =>
        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the most recent fork point */
      case NDCont(cls, argvs, bstore, time, frames)::ndk =>
        val Clos(Lam(v, body), c_env) = cls.head
        val baddr = allocBind(v, time)
        val new_env = c_env + (v -> baddr)
        val new_store = bstore.update(baddr, argvs)
        val new_ndk = NDCont(cls.tail, argvs, bstore, tile, frames)::ndk
        /* resume the fork point with the next closure */
        NDState(body, new_env, new_store, frames, time, new_ndk)
    }
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, newTime)
      val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
      NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
  }
\end{lstlisting}

@@

\begin{lstlisting}
def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val s = nds.toState
      val new_seen = if (seen.contains(s)) seen else seen+s
      val newTime = nds.tick
      val new_ndstate = nds match {
        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) =>
          ...
        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) =>
          ...
      }
      drive_step(new_ndstate, new_seen)
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val s = nds.toState
      val new_seen = if (seen.contains(s)) seen else seen + s
      val new_time = nds.tick
      nds match {
        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
          ...
          drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
        case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
          continue(nds, new_seen)
      }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def continue(nds: NDState, seen: Set[State]): Set[State] = {
  val NDState(ae, env, bstore, konts, time, ndk) = nds
  val new_time = nds.tick
  konts match {
    case Nil => ndcontinue(nds, seen)
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, new_time)
      val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
      drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def ndcontinue(nds: NDState, seen: Set[State]): Set[State] = {
  val NDState(ae, env, bstore, konts, time, ndk) = nds
  ndk match {
    case NDCont(Nil, _, _, _, _)::ndk =>
      drive_step(NDState(ae, env, bstore, konts, time, ndk), seen)
    case NDCont(cls, argvs, bstore, time, frames)::ndk =>
      val Clos(Lam(v, body), c_env) = cls.head
      val baddr = allocBind(v, time)
      val new_env = c_env + (v -> baddr)
      val new_store = bstore.update(baddr, argvs)
      val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
      drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store)
  nd[Storable](closures, Set[VS](), { case (clos, clsnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_cenv = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store))
    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
      nd[VS](bdvss, Set[VS](), { case (vs, bdnd) =>
        val VS(vals, time, store) = vs
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        /* evaluate e with new env and store, then transfer to next body's value */
        aeval(e, new_env, new_store, time, bdnd)
      }, clsnd) /* transfer to next target closure of f */
    })
  },
  continue)
\end{lstlisting}

@@

\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time)
\end{lstlisting}

>>>>

\begin{lstlisting}
type Addr = Int;           type Env = Map[String, Addr];           type Store = Map[Addr, Storable]
abstract class Storable
case class Clos(v: Lam, env: Env) extends Storable
case class Frame(x: String, e: Expr, env: Env);                    type Kont = List[Frame]
case class State(e: Expr, env: Env, store: Store, k: Kont)
\end{lstlisting}

@@

\begin{lstlisting}
case class Frame(x: String, e: Expr, env: Env)
case class Cont(frame: Frame, kaddr: KAddr)
\end{lstlisting}

@@

In the definition of @State@, the continuation store and continuation address disappear;
instead, a list of frames represents the stack. An empty list denotes that we have reached 
the halt. The other components remain unchanged as in the Section \ref{aam}.

@@

\begin{lstlisting}
case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
case class NDState(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time, ndk: List[NDCont]) {
  def toState: State = State(e, env, bstore, konts, time)
  def tick: Time = ...
}
\end{lstlisting}

@@

\begin{lstlisting}
case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
  konts match {
    case Nil => ndk match {
      case NDCont(Nil, _, _, _, _)::ndk =>
        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the most recent fork point */
      case NDCont(cls, argvs, bstore, time, frames)::ndk =>
        val Clos(Lam(v, body), c_env) = cls.head
        val baddr = allocBind(v, time); val new_env = c_env + (v -> baddr)
        val new_store = bstore.update(baddr, argvs)
        val new_ndk = NDCont(cls.tail, argvs, bstore, tile, frames)::ndk
        /* resume the fork point with the next closure */
        NDState(body, new_env, new_store, frames, time, new_ndk)
    }
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, newTime); val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
      NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
  }
\end{lstlisting}

@@

\begin{lstlisting}
def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val s = nds.toState; val newTime = nds.tick
      val new_seen = if (seen.contains(s)) seen else seen+s
      val new_ndstate = nds match {
        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) => ...
        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) => ...
      }
      drive_step(new_ndstate, new_seen)
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val s = nds.toState; val new_time = nds.tick
      val new_seen = if (seen.contains(s)) seen else seen + s
      nds match {
        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
          ...
          drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
        case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
          continue(nds, new_seen)
      }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def continue(nds: NDState, seen: Set[State]): Set[State] = {
  val NDState(ae, env, bstore, konts, time, ndk) = nds; val new_time = nds.tick
  konts match {
    case Nil => ndcontinue(nds, seen)
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, new_time); val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
      drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def ndcontinue(nds: NDState, seen: Set[State]): Set[State] = {
  val NDState(ae, env, bstore, konts, time, ndk) = nds
  ndk match {
    case NDCont(Nil, _, _, _, _)::ndk =>
      drive_step(NDState(ae, env, bstore, konts, time, ndk), seen)
    case NDCont(cls, argvs, bstore, time, frames)::ndk =>
      val Clos(Lam(v, body), c_env) = cls.head
      val baddr = allocBind(v, time); val new_env = c_env + (v -> baddr)
      val new_store = bstore.update(baddr, argvs)
      val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
      drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store)
  nd[Storable](closures, Set[VS](), { case (clos, clsnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time); val new_cenv = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store))
    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
      nd[VS](bdvss, Set[VS](), { case (vs, bdnd) =>
        val VS(vals, time, store) = vs
        val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        /* evaluate e with new env and store, then transfer to next body's value */
        aeval(e, new_env, new_store, time, bdnd)
      }, clsnd) /* transfer to next target closure of f */
    })
  },
  continue)
\end{lstlisting}

@@

\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time) { ... }
\end{lstlisting}
